### 2021-04-23 newvariant.tex
done?

\begin{flushleft}{
	\Large
	\textbf\newline{
		Roles of generation-interval distributions in characterizing relative epidemic strength and speed of new SARS-CoV-2 variants and their control
	}
}
\end{flushleft}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and constant-speed interventions on the spread of the new variant.}
(A--B) Pre-intervention (black) and post-intervention (colored) kernel of the new variant under constant-strength (A) and -speed (B) interventions assuming that the variant has equal generation intervals as the wild type.
(C--D) Pre-intervention (black) and post-intervention kernel of the new variant assuming that the variant has longer generation intervals than the wild type.
(E) Pre-intervention (black) and post-intervention strength conditional on the mean generation interval of the variant (colored).
(F) Pre-intervention (black) and post-intervention speed conditional on the mean generation interval of the variant (colored).
Strength and speed are calculated assuming $\rw=0\pday$, $\delta=0.1\pday$, $\rv=0.1\pday$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
Intervention strength and speed are chosen so that post intervention strength of the new variant is 0.9 when its mean generation interval is 5 days.9.
}
\label{fig:strengthspeed}
\end{figure}

@@

The constant post-intervention speed can be understood by considering a speed-based decomposition of the kernel: $K_{\mathrm{pre}}(\tau) = \exp(\rv) b(\tau)$, where $b(\tau)$ is the initial backward generation-interval distribution and therefore integrates to 1 \citep{doi:10.1098/rspb.2020.1556}.
This decomposition allows us to re-write the post-intervention kernel (under the constant-speed intervention) as follows:
\begin{equation}
K_{\mathrm{post}}(\tau) = \exp(\rv-\phi) b(\tau).
\end{equation}
Since the post-intervention speed $\ry{post}$ is determined by the following implicit equation,
\begin{equation}
1 = \int_0^\infty \exp(-\ry{post} \tau) K_{\mathrm{post}}(\tau) \dtau = \int_0^\infty \exp(\rv-\phi-\ry{post}) b(\tau)  \dtau,
\end{equation}
we have $\ry{post} = \phi - \rv$. In particular, if speed of intervention is faster than the observed speed of spread (i.e., if $\phi > \rv$), we can control the epidemic regardless of the underlying generation-interval distribution.

@@

\jd{If this \P\ is going to be mostly about us, we should say so more clearly.} Since the beginning of the SARS-CoV-2 pandemic, a few studies have emphasized the relevance of leveraging generation interval distributions to improve estimates of strength (e.g., \cite{doi:10.1098/rsif.2020.0144,park2021forward}) and the relative importance of modes of transmission (e.g., asymptomatic vs. symptomatic, \cite{park2020time}).
These same lessons should be considered when assessing the spread of variants in a partially susceptible population and improving efforts to control spread.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions.

>>>>

\begin{flushleft}{
	\Large
	\textbf\newline{
		Roles of generation-interval distributions in characterizing relative epidemic strength and speed of new SARS-CoV-2 variants and their control
	}
}
\\
Sang Woo Park\textsuperscript{1,*}
Benjamin M.\ Bolker\textsuperscript{2,3,4}
Sebastian Funk\textsuperscript{5,6}
C.\ Jessica E.\ Metcalf\textsuperscript{1,7}
Joshua S.\ Weitz\textsuperscript{8,9}
Bryan T.\ Grenfell\textsuperscript{1,7,10}
Jonathan Dushoff\textsuperscript{2,3,4}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Department for Infectious Disease Epidemiology, London School of Hygiene and Tropical Medicine, London, UK
\\
\textbf{6} Centre for Mathematical Modelling of Infectious Diseases, London School of Hygiene and Tropical Medicine, London, UK
\\
\textbf{7} Princeton School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\textbf{8} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{9} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
\\
\textbf{10} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\bigskip

*Corresponding author: swp2@princeton.edu
\bigskip

\end{flushleft}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and constant-speed interventions on the spread of the new variant.}
(A--B) Pre-intervention (black) and post-intervention (colored) kernel of the new variant under constant-strength (A) and -speed (B) interventions assuming that the variant has equal generation intervals as the wild type.
(C--D) Pre-intervention (black) and post-intervention kernel of the new variant assuming that the variant has longer generation intervals than the wild type.
(E) Pre-intervention (black) and post-intervention strength conditional on the mean generation interval of the variant (colored).
(F) Pre-intervention (black) and post-intervention speed conditional on the mean generation interval of the variant (colored).
Strength and speed are calculated assuming $\rw=0\pday$, $\delta=0.1\pday$, $\rv=0.1\pday$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
Intervention strength and speed are chosen so that post intervention strength of the new variant is 0.9 when its mean generation interval is 5 days.
}
\label{fig:strengthspeed}
\end{figure}

@@

The constant post-intervention speed can be understood by considering a speed-based decomposition of the kernel: $K_{\mathrm{pre}}(\tau) = \exp(\rv) b(\tau)$, where $b(\tau)$ is the initial backward generation-interval distribution and therefore integrates to 1 \citep{doi:10.1098/rspb.2020.1556}.
This decomposition allows us to re-write the post-intervention kernel (under the constant-speed intervention) as follows:
\begin{equation}
K_{\mathrm{post}}(\tau) = \exp(\rv-\phi) b(\tau).
\end{equation}
Since the post-intervention speed $\ry{post}$ is determined by the following implicit equation,
\begin{equation}
1 = \int_0^\infty \exp(-\ry{post} \tau) K_{\mathrm{post}}(\tau) \dtau = \int_0^\infty \exp(\rv-\phi-\ry{post}) b(\tau)  \dtau,
\end{equation}
we have $\ry{post} = \phi - \rv$. In particular, if speed of intervention is faster than the observed speed of spread (i.e., if $\phi > \rv$), we can control the epidemic (i.e., $\ry{post} < 0$) regardless of the underlying generation-interval distribution.

@@

Since the beginning of the SARS-CoV-2 pandemic, a few studies have emphasized the relevance of leveraging generation interval distributions to improve estimates of strength \citep{doi:10.1098/rsif.2020.0144,ali2020serial,gostic2020practical,park2021forward} and the relative importance of modes of transmission (e.g., asymptomatic vs. symptomatic, \cite{park2020time}).
These same lessons should be considered when assessing the spread of variants in a partially susceptible population and improving efforts to control spread.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions.
